Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the MIC Architecture

### 2013-10-13 ieee_paper.tex
organizing figures

\newcommand{\todo}[1]{\color{red}\textbf{\hl{#1}}\color{black}\xspace}

@@

\newfig{figures/gflops_peak.pdf}{2.5in}{Peak performance of SpMV
  algorithm assuming either 1 or 4 matrices and/or vectors. We used a
  memory bandwidth of 150 Gbytes/sec (based on measurements in a code
  similar to the SpMV kernel, with computations
  removed.)}{fig:gflops_peak_perf}

\newfig{figures/speedup_wrt_base.pdf}{\wide}{Speedup relative to base
  case using one matrix and one vector.}{fig:speedup}

@@

\newfig{figures/test1_readwrite.png}{\wide}{Bandwidth performance
  under idealized conditions as a function of matrix row size. Entries
  with "cpp" denote cases where coding was performed without MIC
  vector instructions.}{fig:band_rw}

\newfig{figures/test1_gather.png}{\wide}{Bandwidth performance under
  idealized conditions as a function of matrix row size. Entries with
  "cpp" denote cases where coding was performed without MIC vector
  instructions.}{fig:band_gather}

\newfig{figures/test3_gather.png}{\wide}{Bandwidth performance under
  idealized conditions as a function of matrix row size. Entries with
  "cpp" denote cases where coding was performed without MIC vector
  instructions. The greater speed of the cpp version is obtained
  throught the use of \ttt{\#Ivdep} {\em and\/}
  \ttt{\_\_assumed\_aligned}. All memory is aligned on 64
  bytes.}{fig:band_gather_ivdep}

@@

\newfig{figures/host_test1_readwrite_no_temporal_hint.png}{\wide}{Bandwidth
  performance on the host under idealized conditions as a function of
  matrix row size. Entries with "cpp" denote cases where coding was
  performed without AVX vector instructions. The speed on the CPU
  matches the speed with AVX instructions. All memory is aligned on 32
  bytes.}{fig:read_write}

\newfig{figures/mic_performance_nb_threads.png}{\wide}{Performance of
  $y=Ax$ on the MIC. Squares: base 1/1 case, solid circles: 4/4 case
  implemented in C++, solid triangles: 4/4 case implemented with MIC
  vector instructions. $-O3$ compilation options. Each group consists
  of four cases: grids of $64^3$ and $96^3$ with and without Reverse
  Cuthill McKee.}{fig:mic_performance}

% NOTES: why is result the same at 32 processors. I should find out the distribution of threads. 16 threads means one thread on each of 16 cores (YES, I BELIEVE), or 2 threads on each of 8 cores.} Must run the experiment with different nodes orderings. 
%export OMP_SCHEDULE=guided,64    (I should redo experiment with Dynamic or static and compact for KMP_AFFINITY
%export KMP_AFFINITY=compact
% MUST PRINT OUT OMP_SCHEDULE and KMP_AFFINITY inside the program. 
% WHY the very high performance with very few nodes? STRANGE. 
\newfig{figures/host_performance.png}{\wide}{Performance of $y=Ax$ on
  the Host. Squares: base 1/1 case, solid circles: 4/4 case
  implemented in C++, solid triangles: 4/4 case implemented with MIC
  vector instructions. $-O3$ compilation options. The four colors
  distinguish grid resolution and whether or not Reverse Cuthill McKee
  is applied.}{fig:host_performance}

\begin{figure*}[t]
  \centering
  \subfigure[a]{\includegraphics[width=.49\linewidth]{figures/supercompact_max_perf.pdf}}
  \subfigure[b]{\includegraphics[width=.49\linewidth]{figures/compact_max_perf.pdf}}

  \subfigure[c]{\includegraphics[width=.49\linewidth]{figures/random_max_perf.pdf}}
  \subfigure[d]{\includegraphics[width=.49\linewidth]{figures/rbf_max_perf_run1.pdf}}
  \caption{perf accross instance types}
\end{figure*}

@@

>>>>

\newcommand{\todo}[1]{{\color{red}\textbf{\hl{#1}}\xspace}}

@@

\begin{figure*}

  \centering 
  \subfigure[Peak performance of SpMV algorithm assuming either 1 or 4
    matrices and/or vectors. We used a memory bandwidth of 150
    Gbytes/sec (based on measurements in a code similar to the SpMV
    kernel, with computations
    removed).]{\includegraphics[width=.49\linewidth]{figures/gflops_peak.pdf}\label{fig:gflops_peak_perf}}
%
  \subfigure[Speedup relative to base case using one matrix and one
    vector.]{\includegraphics[width=.49\linewidth]{figures/speedup_wrt_base.pdf}\label{fig:speedup}}

  \caption{\todo{to write}}

\end{figure*}

@@

\begin{figure*}

  \subfigure[Bandwidth performance under idealized conditions as a
    function of matrix row size. Entries with "cpp" denote cases where
    coding was performed without MIC vector
    instructions.]{\includegraphics[width=.49\linewidth]{figures/test1_readwrite.png}\label{fig:band_rw}}
%
  \subfigure[Bandwidth performance under
  idealized conditions as a function of matrix row size. Entries with
  "cpp" denote cases where coding was performed without MIC vector
  instructions.]{\includegraphics[width=.49\linewidth]{figures/test1_gather.png}\label{fig:band_gather}}

  \subfigure[Bandwidth performance under idealized conditions as a
    function of matrix row size. Entries with "cpp" denote cases where
    coding was performed without MIC vector instructions. The greater
    speed of the cpp version is obtained throught the use of
    \ttt{\#Ivdep} {\em and\/} \ttt{\_\_assumed\_aligned}. All memory
    is aligned on 64
    bytes.]{\includegraphics[width=.49\linewidth]{figures/test3_gather.png}\label{fig:band_gather_ivdep}}
%
  \subfigure[Bandwidth performance on the host under idealized
    conditions as a function of matrix row size. Entries with "cpp"
    denote cases where coding was performed without AVX vector
    instructions. The speed on the CPU matches the speed with AVX
    instructions. All memory is aligned on 32
    bytes.]{\includegraphics[width=.49\linewidth]{figures/host_test1_readwrite_no_temporal_hint.png}\label{fig:read_write}}

\caption{Bandwidth Analysis.\todo{not 100\% sure what this plots are.}
  \todo{can we harmozie y-axis?} \todo{ylabel should read "Bandwidth (GB/s)".}
  \todo{remove title.}}

\end{figure*}

@@

\begin{figure*}
  \centering

  \subfigure[Performance of $y=Ax$ on the MIC. Squares: base 1/1 case,
    solid circles: 4/4 case implemented in C++, solid triangles: 4/4
    case implemented with MIC vector instructions. $-O3$ compilation
    options. Each group consists of four cases: grids of $64^3$ and
    $96^3$ with and without Reverse Cuthill
    McKee.\todo{I don't think this is supercompact but RBF 3D.}]{\includegraphics[width=.49\linewidth]{figures/mic_performance_nb_threads.png}\label{}}
%
  \subfigure[performance on the host on a 64x64x64 matrix. Manual
    vectoization brings significant improvement there as
    well.]{\includegraphics[width=.49\linewidth]{figures/64x64x64.pdf}\label{}}

  \caption{Vectorization matters in practice. Differences in BW
    translate in different performance.\todo{this two plots should ``correlate'' }}
\end{figure*}

\begin{figure*}[t]
  \centering
  \subfigure[Supercompact\todo{is the title? this looks like super compact.}]{\includegraphics[width=.49\linewidth]{figures/supercompact_max_perf.pdf}}
  \subfigure[Compact]{\includegraphics[width=.49\linewidth]{figures/compact_max_perf.pdf}}

  \subfigure[Random]{\includegraphics[width=.49\linewidth]{figures/random_max_perf.pdf}}
  \subfigure[RBF 3D]{\includegraphics[width=.49\linewidth]{figures/rbf_max_perf_run1.pdf}}
  \caption{Performance of "best code" accross instance types. \todo{Why stop at 1.6M
      rows. Is that because of memory limitation? I guess that's what
      it is}\todo{can we remove the title?}\todo{can we harmonize y-axis?}}
\end{figure*}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/plot_for_natasha_ncar_end_of_year.pdf}

  \caption{Performance on a MIC of derivative computation for RBFFD
    with 884,736 nodes distributed quasi-randomly in a cube. Single
    derivative of a single function (base case, triangle) and four
    derivatives of four functions (multi case, circles). No bandwidth
    reduction (blue) and Reverse Cuthill-McKee (red).\todo{should we have a similar one for the host?}}
\end{figure}

Derivatives in an RBFFD formulation are expressed as a sparse
matrix/vector multiplication (SpMV). Using the full 61 nodes of a MIC
processor, and all four threads of each node, we achieve lowly 18
Gflops when calculating a single vector of a single
function. Bandwidth reduction has minimal effect since the cache is
large enough to store both the derivative matrix and the solution
vector. We increase the number of computations per byte transferred by
calculating four different derivatives of four different
functions. With bandwidth reduction, we speed the calculation by a
factor of 8, and achieve upwards of 140 Gflops.

@@

future works: gpu. more applications. more nnz/row.
